---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo=TRUE}
library(dplyr)
library(ggplot2)

# Read the dataset
activity <- read.csv("activity.csv")

# Check the structure
str(activity)
```


## What is mean total number of steps taken per day?
```{r echo = TRUE}
# Calculate total steps per day
total_steps <- activity %>%
  group_by(date) %>%
  summarize(total = sum(steps, na.rm = TRUE))

# Plot histogram
hist(total_steps$total, breaks=20, main="Total Steps per Day", xlab="Steps", col="blue")

# Mean and Median
mean_steps <- mean(total_steps$total)
median_steps <- median(total_steps$total)

mean_steps
median_steps
```

## What is the average daily activity pattern?
```{r echo = TRUE}
# Average steps per interval
avg_steps_interval <- activity %>%
  group_by(interval) %>%
  summarize(mean_steps = mean(steps, na.rm = TRUE))

# Time series plot
plot(avg_steps_interval$interval, avg_steps_interval$mean_steps, type = "l",
     xlab="5-minute Interval", ylab="Average Steps", main="Average Daily Activity Pattern")

# Interval with maximum average steps
max_interval <- avg_steps_interval[which.max(avg_steps_interval$mean_steps), ]
max_interval
```


## Imputing missing values

```{r echo = TRUE}
# Total number of missing values
sum(is.na(activity$steps))

# Fill NA with mean for that interval
activity_imputed <- activity
activity_imputed$steps[is.na(activity_imputed$steps)] <- 
  avg_steps_interval$mean_steps[match(activity_imputed$interval, avg_steps_interval$interval)]

# New total steps per day
total_steps_imputed <- activity_imputed %>%
  group_by(date) %>%
  summarize(total = sum(steps))

# Plot histogram after imputation
hist(total_steps_imputed$total, breaks=20, main="Total Steps per Day (Imputed)", xlab="Steps", col="green")

# New mean and median
mean_steps_imputed <- mean(total_steps_imputed$total)
median_steps_imputed <- median(total_steps_imputed$total)

mean_steps_imputed
median_steps_imputed
```

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
# Add a new column: weekday or weekend
activity_imputed$day_type <- ifelse(weekdays(as.Date(activity_imputed$date)) %in% c("Saturday", "Sunday"), 
                                    "weekend", "weekday")

# Average steps by interval and day type
avg_steps_daytype <- activity_imputed %>%
  group_by(interval, day_type) %>%
  summarize(mean_steps = mean(steps))

# Plot
library(lattice)
xyplot(mean_steps ~ interval | day_type, data=avg_steps_daytype, type="l",
       layout=c(1,2), xlab="Interval", ylab="Number of steps")

```

## Conclusion
After imputing missing values, mean and median were slightly different.

Weekends and weekdays show different step patterns.
